### 2024-09-11 paper-en.tex
Update translation

    Suppose $\eta' = (e_1, ..., e_{d - 1})$ is a path. Let us call $x$ the last vertex of path $\eta$. According to the supposition of induction the last vertices of it's corresponding paths on extension are $x^{(i + \chi(\eta')(\mod{r}))}$.

@@

\begin{theorem}[on cycles]\label{theorem_cycles}
    Let $\eta$ be a cycle on metagaph. Then it's corresponding paths on the extension are cycles iff $\chi(\eta) = 0 (\mod{r})$.
\end{theorem}

\begin{theorem}[on wings of cycles]\label{theorem_wings}
    Let $\eta = (e_1, ..., e_d)$ and $\mu = (\epsilon_1, ..., \epsilon_{\delta})$ be different paths on metagraph $G$, with first vertex $a$, and last vertex $b$, $\chi(\eta) = \chi(\mu) (\mod{r})$, and let $\gamma = (e_1, \dots, e_d, \epsilon_{\delta}, \dots, \epsilon_1)$ be a path on $G$. Then paths on extension which correspond to $\gamma$, are cycles.
\end{theorem}

@@

>>>>

    Suppose $\eta' = (e_1, ..., e_{d - 1})$ is a path. Let us call $x$ the last vertex of path $\eta$. According to the supposition of induction the last vertices of it's corresponding paths on expansion are $x^{(i + \chi(\eta')(\mod{r}))}$.

@@

\begin{theorem}[on cycles]\label{theorem_cycles}
    Let $\eta$ be a cycle on metagaph. Then it's corresponding paths on the expansion are cycles iff $\chi(\eta) = 0 (\mod{r})$.
\end{theorem}

\begin{theorem}[on wings of cycles]\label{theorem_wings}
    Let $\eta = (e_1, ..., e_d)$ and $\mu = (\epsilon_1, ..., \epsilon_{\delta})$ be different paths on metagraph $G$, with first vertex $a$, and last vertex $b$, $\chi(\eta) = \chi(\mu) (\mod{r})$, and let $\gamma = (e_1, \dots, e_d, \epsilon_{\delta}, \dots, \epsilon_1)$ be a path on $G$. Then paths on expansion which correspond to $\gamma$, are cycles.
\end{theorem}

\section{On graphs with given girth.}

% help us __to__ build?

Let $G(V, E, f, w)$ be a metagraph and $w(e_i) = x_i \; \forall i = 0, \dots, |E|$, where $x_i$ are variables that will help us build a graph with given girth.

% Поискать, как в англоязычной литературе называют не равенства с \neq

Since we have to build an expansion graph $G'$ with girth $g$, therefore for any cycle $\eta$ with length less than $g$ on metagraph $G$, assume inequality of the form $\chi(\eta) \neq 0$.

\begin{theorem}

Let $G$ be a metagraph. Then there exists an expansion of metagraph $G$ with girth equal to or greater than $g$ iff the following system of inequalities for a graph $G$ has a solution:

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            \chi(\eta) &\neq 0 (\mod{r}) \; \forall \, \eta \in \mathcal{H}_G(g).
        \end{array}
    \right.,
    \label{eqs:example}
\end{equation}
where $\mathcal{H}_G(g)$ is the set of all cycles on $G$ with length less then $g$.

\begin{proof}
    % exhausts?
    The construction of system (1) exhausts all cycles, hence the absence of a solution is equivalent to existence of a cycle with length less then $g$ on any expansion. Thus the existence of a solution for system (1) is equivalent to existence of an expansion with girth greater than $g$.
\end{proof}

@@

% Добавить везде где надо (\mod{r})

For a given metagraph $G$, system \eqref{eqs:example} can be built using the following algorithm.

\textbf{Algorithm 1.}

Let $G(A \cup B, E, f, w)$ be a metagraph and let $w(e_i) = x_i \; \forall i = 0..|E|$.

% Нужно заменить g на какую-то не используемую выше букву и заменить во всей теореме

This algorithm is based on the construction of the set of labels.
Each such label is a quadruple $(v, l, ch, p)$,
$v$ is a vertex marked by this label;
where $l$ is a length of the path $\mu$ from starting vertex to the vertex $v$; 
$ch \in \mathbb{Z}$ characteristic of the path $\mu$;
$p \in V \cup \{ nil \}$ is an edge that was used to build this label, $nil \not\in V $ is a special value reserved for a starting label.

Assign the system of all inequalities $\mathcal{S}$ to be empty.
For each vertex $a \in A$:

\begin{itemize}
    \item Add starting label $(a, 0, 0, nil)$ to an empty set of all labels,
    \item Iterating by $l$ from $1$ to $g / 2$:
    \begin{itemize}
        \item For all labels $(v, l - 1, ch, p)$:
        \item
            For each edge $e \not= p$ incident to vertex $v$,
            build label $(v', l, ch', p)$, where $ch' = ch + (-1)^{l} w(p)$, $v'$ is a vertex different from $v$, incident to $p$.
            Add this label to the set of all labels.
    \end{itemize}
    \item for all pairs of labels $(v_1, l_1, ch_1, p_1)$ and $(v_2, l_2, ch_2, p_2)$ such that $v_1 = v_2$ and $l_1 = l_2$, add inequality $ch_1 \neq \ch_2$ to the system $\mathcal{S}$.
\end{itemize}

\begin{notice}
    Algorithm 1 builds cycles with two similar "wings", i.e. cycles that consist of two path of equal length. All cycles on bipartite graph can be represented in such a way, thus all cycles from $\mathcal{H}_G(g)$ on metagraph will be found by Algorithm 1.
\end{notice}

\textbf{Example.}

In figure \ref{neq_system_graph} the metagraph $G$ with for edges $e_1, e_2, e_3, e_4$ is shown. Weights of edges are variables $x_1, x_2, x_3, x_4$.

\begin{figure}[H]
    \centering
    \begin{picture}(150,200)
        \put(75,165){\thicklines{\circle*{5}}}
        \put(70,170){$v_1$}

        \put(35,35){\thicklines{\circle*{5}}}
        \put(30,20){$v_2$}

        \put(115,35){\thicklines{\circle*{5}}}
        \put(110,20){$v_3$}

        \bezier{300}(75,165)(10,100)(35,35)
        \put(5,100){$x_1$}

        \bezier{300}(75,165)(56,100)(35,35)
        \put(56,80){$x_2$}

        \bezier{300}(75,165)(140,100)(115,35)
        \put(125,100){$x_3$}

        \bezier{300}(75,165)(94,100)(115,35)
        \put(87,80){$x_4$}
    \end{picture}
    \caption{ Metagraph $G$, with variable weights of edges $x_1, x_2, x_3, x_4$. }
    \label{neq_system_graph}
\end{figure}

% "girth to be 6." ???
Let us apply Algorithm 1 for metagraph $G$ and girth to be 6. The set of labels, constructed by the Algorithm is shown in Table \ref{cycle_search_table_neq}. The system of inequalities (2) is constructed using these labels.

\begin{table}[H]
    \centering
    \begin{tabular}{ | c | c | c | }
        \hline
        $g = 0$            & $g = 1$               & $g = 2$                   \\ \hline
        $(0, v_1, 0, nil)$ & $(1, v_2,  x_1, e_1)$ & $(2, v_1,  x_1 - x_2, b)$ \\ \hline
                           & $(1, v_2,  x_2, e_2)$ & $(2, v_1,  x_2 - x_1, a)$ \\ \hline
                           & $(1, v_3,  x_3, e_3)$ & $(2, v_1,  x_3 - x_4, e)$ \\ \hline
                           & $(1, v_3,  x_4, e_4)$ & $(2, v_1,  x_4 - x_3, d)$ \\ \hline
    \end{tabular}
    \caption{ The set of labels constructed during the execution of the Algorithm 1. }
    \label{cycle_search_table_neq}
\end{table}

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            x_1 - x_2 &\neq 0             ; \\
            x_3 - x_4 &\neq 0             ; \\
            2 x_1 - 2 x_2 &\neq 0         ; \\
            2 x_3 - 2 x_4 &\neq 0         ; \\
            x_1 - x_2 - x_3 + x_4 &\neq 0 ; \\
            x_1 - x_2 + x_3 - x_4 &\neq 0 . \\
        \end{array}
        \label{eqs:cycle_search_neqs}
    \right.
\end{equation}

\section{Solving systems of inequalities}

% Абзац про минимизацию r

% Still ???

Finding of the minimal solution for system \eqref{eqs:example}, is a computationally hard problem. It requires to enumerate all possible solutions and select the minimal one. Still, if system \eqref{eqs:example} has a solution then one of them can be found by the following algorithm.

\textbf{Algorithm 2.}

\begin{enumerate}
    \item Let $e$ be the number of variables in a system of inequalities.
    \item If the system contains inequality $0 \neq 0$ then it has no solution, the algorithm halts.
    \item Let us pick all inequalities of the form $a x_e + b \neq 0$. Let $I$ be the set of all such inequalities.
    \item Let $v$ be a number such that $v \neq -b/a \forall a x_e + b \in I$, if $I = \varnothing$, then let $v = 0$. Такое число можно найти, так как не равенств конечное количество. В результате этой замены не возникнет не равенства вида $0 \neq 0$ по условию выбора $v$.
    \item Заменим во всех не равенствах $x_e$ на $v$ и добавим в решение элемент $x_e = v$.
    \item Если $e = 1$ ~-- завершим работу алгоритма.
    \item Положим $e = e - 1$ переменной и вернёмся на шаг 3.
\end{enumerate}

\textbf{Пример.}

Положим $x_1 = 0$, подставим в систему не равенств, получим:

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - x_2 &\neq 0            ; \\
            x_3 - x_4 &\neq 0        ; \\
            - 2 x_2 &\neq 0          ; \\
            - 2 x_4 &\neq 0          ; \\
            - x_2 - x_3 + x_4 &\neq 0; \\
            - x_2 + x_3 - x_4 &\neq 0. \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs}
\end{equation}

Значение $x_2$ не может быть равно $0$, т.к. в таком случае нарушается не равенство $- x_2 \neq 0$. Положим $x_2 = 1$, подставим в систему не равенств, получим:

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0             ; \\
            x_3 - x_4 &\neq 0       ; \\
            - 2 &\neq 0             ; \\
            - 2 x_4 &\neq 0         ; \\
            - 1 - x_3 + x_4 &\neq 0 ; \\
            - 1 + x_3 - x_4 &\neq 0 . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs}
\end{equation}

Положим $x_3 = 0$, подставим в систему не равенств, получим:

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0       ; \\
            - x_4 &\neq 0     ; \\
            - 2 &\neq 0       ; \\
            - 2 x_4 &\neq 0   ; \\
            - 1 + x_4 &\neq 0 ; \\
            - 1 - x_4 &\neq 0 . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs}
\end{equation}

Значение $x_4$ не может принимать значения $0, 1, -1$, т.к. в таком случае нарушаются не равенства. Положим $x_4 = 2$, подставим в систему не равенств, получим:

\begin{equation}
    \centering
    \left\{
        \begin{array}{ll}
            - 1 &\neq 0   ; \\
            - 2 &\neq 0   ; \\
            - 2 &\neq 0   ; \\
            - 4 &\neq 0   ; \\
            1 &\neq 0     ; \\
            - 3 &\neq 0   . \\
        \end{array}
    \right..
    \label{eqs:cycle_search_neqs}
\end{equation}

Полученные не равенства показывают, что $r$ не может принимать значения $1, 2, 3, 4$, следовательно минимально допустимое значение $r$ ~-- $5$. 5-расширение метаграфа с весами, полученными в ходе решения системы не равенств, изображено на рисунке \ref{neq_system_res}. Заметим, что действительно обхват этого графа равен 6, один из циклов-шестёрок выделен жирным.

\begin{figure}[H]
    \centering
    \begin{picture}(380,200)
        \put(50,165){\thicklines{\circle*{5}}}
        \put(35,35){\thicklines{\circle*{5}}}
        \put(65,35){\thicklines{\circle*{5}}}

        \bezier{300}(50,165)(42,100)(35,35)
        \bezier{300}(50,165)(57,100)(65,35)

        \put(120,165){\thicklines{\circle*{5}}}
        \put(105,35){\thicklines{\circle*{5}}}
        \put(135,35){\thicklines{\circle*{5}}}

        \bezier{300}(120,165)(127,100)(135,35)
        \bezier{300}(120,165)(197,100)(275,35)

        \put(190,165){\thicklines{\circle*{5}}}
        \put(175,35){\thicklines{\circle*{5}}}
        \put(205,35){\thicklines{\circle*{5}}}

        \bezier{300}(190,165)(217,100)(245,35)
        \bezier{300}(190,165)(267,100)(345,35)

        \put(260,165){\thicklines{\circle*{5}}}
        \put(245,35){\thicklines{\circle*{5}}}
        \put(275,35){\thicklines{\circle*{5}}}

        \bezier{300}(260,165)(252,100)(245,35)
        \bezier{300}(260,165)(267,100)(275,35)
        \bezier{300}(260,165)(287,100)(315,35)
        \bezier{600}(260,165)(172,50)(65,35)

        \put(330,165){\thicklines{\circle*{5}}}
        \put(315,35){\thicklines{\circle*{5}}}
        \put(345,35){\thicklines{\circle*{5}}}

        \bezier{300}(330,165)(322,100)(315,35)
        \bezier{300}(330,165)(337,100)(345,35)
        \bezier{600}(330,165)(182,150)(35,35)
        \bezier{600}(330,165)(242,50)(135,35)

        \thicklines
        \bezier{300}(50,165)(77,100)(105,35)
        \bezier{300}(120,165)(112,100)(105,35)
        \bezier{300}(120,165)(147,100)(175,35)
        \bezier{300}(190,165)(182,100)(175,35)
        \bezier{300}(190,165)(197,100)(205,35)
        \bezier{300}(50,165)(127,100)(205,35)
    \end{picture}
    \caption{ Расширение метаграфа с весами дуг, полученными решением системы не равенств. }
    \label{neq_system_res}
\end{figure}

\begin{notice}
    Результатом работы алгоритма поиска решения системы не равенств не обязательно будет минимальное решение.
\end{notice}
